Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-07-30 documents/ovando-etal-assessing-global-fisheries.Rmd
1st real draft of paper done

---
title: "The Future of Global Fishery Assessments"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals. Numerous high-profile publications have produced estimates of the global status of fisheries, but limited quantity and quality of data, conflicting results, and methodological concerns mean that our understanding of global fisheries is still highly uncertain, especially in critical areas such as Southeast Asia and Africa. We present an evaluation of the past and future of global fishery assessments. Global assessments have typically involved aggregating best-available stock assessments where possible with various data-limited assessments for regions lacking formal evaluations of stock status. The data-limited methods used have ranged from simple catch heuristics to super-ensemble modeling. Despite attempts at model improvement, there have been discrepancies both among global fishery assessments and between the data-limited components of these methods and more data-rich models,demonstrating that our understanding of global data-limited fisheries remains limited. How then can we most efficiently move forward? We utilize a flexible assessment model together with the RAM Legacy Stock Assessment Database not strictly as a stock assessment tool, but to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion and increasingly available but historically underutilized data such as trawl footprints. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world. We close by discussing priority areas for future data collection. Our results make the case that improving our understanding of global fisheries depends not on better models, but on better data.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(here)
extrafont::loadfonts()

@@

results_dir <- here("results",results_name)

load(file = file.path(results_dir, "paper_plots.Rdata"))

pub_theme <- theme_ipsum(base_size = 12, 
                         axis_text_size = 12) + 
    theme(panel.spacing = unit(1,"lines"))

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fishery, 6.2% are underfished ,and 34.2% are overfished [@fao2020].

However, there is room for substantial improvement in the methods used in estimating global stock status. 49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.

The SOFIA report bases its estimates of included unassessed stocks mostly on expert elicitation. While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012;@costello2016a;@rosenberg2018;@pauly2007].Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models [@free2020 and references therein]. These models seek to infer stock status, for example in terms of biomass *b* relative to the biomass at maximum sustainable yield *b~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, given that a global assessment by its nature requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks (high variance, low bias).Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.

However, @free2020 demonstrated that these catch-only models can often produce both imprecise and biased estimates of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but based on surveys and expert opinion we believe it should be worse than Northeast Pacific. The Costello et al. 2016 [@costello2016a] model shows the correct general pattern in terms of relative ranking of regions, but we would expect the state of fisheries in the Mediterranean/Black sea regions and SE Asia to be much worse, and the NE Pacific better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  knitr::kable(caption = "Status by FAO Statistical area from four different assessment methods",
               booktabs = TRUE) %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

@@

We use a new assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of information on global fisheries. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@anderson2015], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model. `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.

@@

Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status. For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region. We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martel2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index. By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic). Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17. Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).

@@

If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense. However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal. Supplementing these data with a local index of abundance provides a more nuanced figure, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar"). While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.

@@

It should be no surprise that incorporating locally tailored data produces more locally plausible results. Global assessment at the moment though requires depending on a few broadly available datasets. To give an illustration, we used `sraplus` to assess the status of in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from @rousseau2019 (xx technically this is the updated dataset xx). Catch histories for Europe and Northeast Asia look somewhat similar xx plot of catch and effort histories xx. As a result, using a "catch-only" heuristic, both Europ and Northeast Asia are estimated to have very similar stock status, a similar finding as those illustrated in Table.XX. Adding in the effort data (as an effort index created for each stock as catch divided by the effective harvest rate based on the effort data), and the two regions begin to diverge. While according to @rousseau2019, Europe has seen a recent decrease in fishing effort from a peak in XX, effort in Northeast Asia has increased more or less continuosly. As a result, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe. However, the shared effort series across these stocks aslo produces a large number of effectively unfished stocks; a highly unlikely result for both regions.

XX need to redo this analyis with the new effort data XX

```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019. Each point is a stock, point size is a function of stock size."}
continent_kobe_plot
```

@@

This exercise reveals the potential and challenges using broadly available data for global fishery assessment. On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results. However, both results also raise a host of new issues. How realistic is it to share one index of abundance and SAR value for the entire eastern Bay of Bengal? Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished (a dynamic which does not match assessments from that region XX plot RAM stocks for europe XX).

Extending these issues to the global level, we assessed the status of stocks assessed by the FAO's latest SOFIA report using various types of broadly available data. Given that the SOFIA assessments present the best estimates of stock status across both assessed and unassessed stocks, we would hope that any assessment method suitable would produce broadly match the values reported by the FAO. We focus here on the percent of stocks classified as "overfished" (B/B~MSY~ < 0.8) in each FAO statistical region. Following the SOFIA methodology, "Stocks" are classified by the sum catch histories across the same species within a major FAO statistical region.

Our perception of global fishery depends completely on which data sources we utilize. That different data tell different stories should not be surprising. However, we also do not see a convergence towards a cohesive picture of global status that closely matches the SOFIA report as we add more data. The most data-poor method employed, the catch heuristic alone, matches some patterns from the SOFIA report (little overfishing in the Northeast Pacific, some overfishing in the northwest Atlantic), it misses the mark badly in other regions, e.g. the waters around New Zealand. Our most data-rich scenario, an effort series from @rousseau2019, along with priors from swept-area-reatio and fisheries management index where available....

to provide more plausible seeming results than catch data alone. A reasonable question then is, can we simply augment global catch statistics with the data presented here to provide an improved understanding of global fisheries?

This is the main plot of the paper Percent of stocks classified as overfished depending on what data you use. CPUE data are at the continental scale (e.g. use North America Rousseau effort for USA, Mexico, Canada).

key message: Different data sources give you different estimates of stock status, and not in a way that makes it clear that any one is inherently superior to another

A few key questions here

- Currently this runs sraplus on all stocks in the region. I think this is informative since it says, suppose we only used sraplus, what would we think, and can compare that to RAM etc. in our heads (or could add a map panel). An alternative though would be to stitch together status, i.e. use RAM where there is RAM, use sraplus otherwise. That would be a better "estimate" of the actual stock status in an area, but would make the contrast between the methods less clear

- This used the continent level Rousseau data - can we get the finer scale version for this paper? Would make things a lot clearer I think

*will add in RAM panel once I match up RAM stocks to FAO region*

```{r po-map, fig.width=8, fig.height=8, fig.cap="Estimated percent of fisheries 'overfished' (B/B_msy_ less than 0.8) using different dat sources. CPUE refers to an index of abundance based on the Rousseau effort data. Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data. The FAO Report 569 is our reference status estimate provided by the FAO. ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region. fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
po_map_plot
```

## Value of Information

Provide some guidance on where to prioritize data collection. e.g. how long of a CPUE series is needed, expansion of f based priors / value of f, advocate for merging together of the "Catch only" world with the "length-based world". E.g. we talk about catch being the only widely available statistic, but in many data-limited communities all they have are lengths. A global effort to create a centralized database of length distributions could allow for length-based priors on spr/f that could be paired with catches?

>>>>

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results and therefore, high uncertainties in our understanding of global fisheries. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a new flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(viridis)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(here)
extrafont::loadfonts()

functions <- list.files(here::here("functions"))

purrr::walk(functions, ~ source(here::here("functions", .x)))

@@

results_path <- here("results",results_name)

min_years_catch <- 25

prepare_sofia_data(lookup_fmi_names = FALSE)

ram_comp_data <- ram_data %>%
  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
  filter(has_things) %>%
  group_by(stockid) %>%
  mutate(delta_year = as.integer(year - lag(year))) %>%
  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
                                TRUE ~ delta_year)) %>%
  mutate(missing_gaps = any(delta_year > 1)) %>%
  filter(missing_gaps == FALSE) %>%
  group_by(stockid) %>%
  mutate(n = length(catch)) %>%
  filter(n >= min_years_catch) %>%
  ungroup()

ram_comp_data <- ram_comp_data %>%
  select(
    year,
    stockid,
    country,
    primary_FAOarea,
    scientificname,
    commonname,
    year,
    catch,
    b_v_bmsy,
    u_v_umsy,
    total_biomass
  ) %>%
  rename(scientific_name = scientificname) %>% 
  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
  left_join(fao_species, by = "scientific_name")

assess_ram_fits <- read_rds(path = file.path(results_path,"assess_ram_fits.rds"))

ram_v_sraplus_area <- read_rds(path = file.path(results_path,"ram_v_sraplus_area.rds"))

ram_comp_data <- read_rds(path = file.path(results_path,"ram_comp_data.rds"))

fao_sraplus_comp_data <- read_rds(path = file.path(results_path, "fao_sraplus_comp_data.rds"))

load(file = file.path(results_path, "paper_plots.Rdata"))

load(file = file.path(results_path, "voi_fits.RData"))

pub_theme <- theme_ipsum(base_size = 10,
                         axis_text_size = 10,
                         axis_title_size = 12) +
    theme(panel.spacing = unit(.5,"lines"))

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fishery, 6.2% are underfished ,and 34.2% are overfished [@fao2020]. While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment. Although revised periodically over time, the method has kept revisions to a minimally necessary levels for the purposes of temporal consistency and comparability. With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert elicitation. While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].

Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (@free2020 and references therein). These models seek to infer stock status, for example in terms of biomass *b* relative to the biomass at maximum sustainable yield *b~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, given that a global assessment by its nature requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks (high variance, low bias).Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.

However, @free2020 demonstrated that these catch-only models can often produce both imprecise and biased estimates of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but other data sources such as surveys and expert opinion suggest this region should have more depleted fisheries than Northeast Pacific. The Costello et al. 20162 model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  knitr::kable(caption = "Status by FAO Statistical area from four different assessment methods",
               booktabs = TRUE) %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@melnychuk2017], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model. `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.

@@

Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status. For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region. We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martell2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index. By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic). Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17. Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).

@@

If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense. However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal. Supplementing these data with a local index of abundance provides a more nuanced result, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar"). While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.

@@

It should be no surprise that incorporating locally tailored data produces more locally plausible results. Global assessment at the moment though depend on few broadly available datasets. We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx). Catch histories for Europe show a steady increase reaching a peak in the 1990s, and a decline and stabilization since. Current catches in Europe are much higher than those at the start of the catch history. Catches in northeast Asia have steadily increased over time, with some stabilization as of late. For both regions, recent catches expressed as a fraction of maximum catch, they key metric from the perspective of a catch-only heuristic, are both high. However, the two regions exhibit markedly different effort histories according to @rousseau2019, with effort in Europe peaking with catches and subsequently declining, while effort in northeast Asia has increased steadily over time. As a result, using a "catch-only" heuristic, both Europe and Northeast Asia are estimated to have very similar stock status. Adding in an index of abundance, calculated for each stock as catch divided by the effective harvest rate based on the effort data for the region, produces a very different picture of stock status and introduces some contrast between the two regions. Based on the catch and effort data, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe. However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).

```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019 Each point is a stock, point size is a function of stock size."}
continent_kobe_plot
```

@@

```{r perf-tab}
breaks <- c(0.8, 1.2)

breaks <- c(0, breaks, Inf)

labels <- c("over", "fully", "under")

rand_fits <- assess_ram_fits %>% 
  filter(data == "ram-data") %>% 
  mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
         data = "guess")

perf <- assess_ram_fits %>% 
  bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
  group_by(data) %>% 
  summarise(
    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
  ) %>% 
  arrange((mape))

knitr::kable(perf, digits = 2,
             caption = "Global summary statistics of models using different sources of data. mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).",
              format = "pipe")

These case studies reveals the potential and challenges of using broadly available data for global fishery assessment. On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results in some ways. However, both case studies also raise a host of new issues. How realistic is it to share one index of abundance and SAR value for the entire eastern Bay of Bengal? Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished.

Extending these issues to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM. Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries. As a proof of concept, we first fit models to abundance indices from each stock in RAM. There are two main reasons why an assessment model might produce poor results: model misspecification and poor data. Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good. The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias. The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time. Looking across space, we see clear geographic patterns in model performance. The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada). The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.

Performance limitations then are likely to arise less from model misspecification but from the quality of the data themselves. These becomes even clearer once we consider the performance of `sraplus` models fit to globally available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1. This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.

Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process. We also see that across the models performance (or lack thereof) is not consistent in space: different data to best/worst for different FAO regions. For example, the catch heuristic and models fit to nominal CPUE data both substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and none of the datasets were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.

```{r dat-description}
dat_description <- tribble(~`Data Name`, ~Description,
        "RAM-Data", "Fit to abundance index from RAM",
        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
        "catch-only", "priors on terminal B/Bmsy informed by catch history",
        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")

knitr::kable(dat_description,
             caption = "Data sources used for terminal stock status estimate",
             format = "pipe")
# %>%
#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
#   kableExtra::row_spec(0, bold = T)
```

```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources. ram-data refers to catch and abundance index drawn from RAM. CPUE refers to an index of abundance based on the Rousseau effort data. Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data. The FAO Report 569 is our reference status estimate provided by the FAO. ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region. fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
ram_mpe_map_plot
```

```{r mape-map, fig.cap = "Median absolute percent error in most recent B/B~MSY~ by FAO statistical area from different data sources. ram-data refers to catch and abundance index drawn from RAM. CPUE refers to an index of abundance based on the Rousseau effort data. Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data. The FAO Report 569 is our reference status estimate provided by the FAO. ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region. fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
ram_mape_map_plot
```
